% -*- mode: latex; coding: utf-8; fill-column: 72; -*-

\footnote{We are still working on the name of this hypothesis, as while the spirit of the hypothesis is similar to the pragmatic syntactic bootstrapping hypothesis proposed by \textcite{hacquardlidz2018} to address the learning problem of attitude verb categories, they seem to differ in many ways. In \textcite{hacquardlidz2018}, the hypothesis is proposed for the learning of an abstract semantic property (the semantic class of an attitude verb) that is systematically related to (a) syntactic distribution and (b) pragmatic function. The \hypos{} is proposed to deal with the problem that observations of only the syntactic distribution or only the pragmatic function had the potential to mislead, so each was used to constrain the other. In the current problem of learning clause types and speech acts, the learners need to figure out two abstract categories, the clause types (abstract formal category) and the speech acts information (abstract semantic/pragmatic category), and the hypothesis is that two categories need to be learned in tandem and mutually constraining ways because formal features and social pragmatic features alone are not sufficient for the learning of clause types and speech acts respectively. 

But we could say that what in fact needs to be learned in our situation, is a semantic category underlying the clause type category and speech act category, namely \tit{mood} or \tit{sentence force}. This category is systematically related to (a) syntactic distribution (i.e. the distribution of formal features that we are currently assuming to be associated with clause type features) and (b) pragmatic functions (speech acts). Observations of only the formal distribution or only the pragmatic function had the potential to mislead, and so the learners need both to mutually constrain each other. In this way, the two versions of \hypos{} might share more similarities.




When examined separately, the learning of clause types and the learning of speech acts share many similarities. Clause type is an abstract formal feature of a sentence related to a variety of surface forms, none of which is obligatorily present and many of which can occur in sentences with a different clause-type feature. While in principle the surface formal features could be sufficient, this dissertation showed that they in fact fall short. To compensate for this insufficiency, learners can use the speech act information -- which is systematically related to clause types -- to learn to identify clause types and which surface features are relevant for clause typing. Speech act is also an abstract category. As we have seen in Chapter~\ref{chap:eng-sp}, speech act categories are related to a variety of human behaviors (e.g. pauses and eye gaze), none of which are obligatorily present (you don't \emph{need} to pause after a question) and many of which can occur when performing other speech act categories. While I showed that the features are correlated with the use of questions, it is likely that the learners use the clause type information --- which is systematically related to speech acts --- to inform how they infer speech act types.


%%prosody
Comment about sec. 5.2.3.1 (on prosody): I think you can pitch this in a much more positive light. Just tell the reader what you are interested in (nuclear contours and whether they rise or fall), and what that means (the nuclear contour is the contour from the nuclear pitch accent to the end of the utterance; the nuclear pitch accent is the final pitch accent, and a pitch accent is just a marker of greater relative prominence on a syllable; and finally, prominence is a perceptual category: a syllable that is prominent is in some sense intuitively "stronger" than surrounding syllables, according to human intuition; you can cite the readings for the ToBI course here https://ocw.mit.edu/courses/6-911-transcribing-prosodic-structure-of-spoken-utterances-with-tobi-january-iap-2006/, and also Büring's 2016 book, and let me know if you don't have it). Then you say you want to look at thousands of sentences, and there's no way to automate recognition of (nuclear) pitch accents (you can cite Büring 2016). Ultimately you hope to get a hand annotation, but for the moment you came up with this algorithm for automating the task by looking at the final peak or valley and what the contour does afterward. You compared it to a hand annotation of a small subset, and it's not a perfect measure, but it's on the right track. You might also note here that one of the big problems (besides the fact that the recordings themselves are literally noisy) is that you can't trust the tail end of all pitch tracks. And in fact the one example you gave in fig. 5.2 looks to me like an example of this (see my notes in the paper copy). The reason is that pitch periods are irregular due to dissipating energy. This can cause spurious events in the waveform to be misinterpreted as a pitch period, which Praat then incorrectly interprets as pitch. See a brief discussion on this in chapter 2.4 of the ToBI course. Finally, you give the plots you already have in this subsection, and you say, despite being a noisy approximation of the actual data, you found some signal where it matters (polar interrogatives have more rises!). And so the child might be using this, and you say that you plan to look at higher quality hand annotated data in future work. (You don't need to put this in, but I think you might also hope to improve the automated categorization as well.)

And then my main question: To summarize, what do you take the role of prosody to be in resolving these acquisition puzzles? How does the answer to that guide the way you use it in your studies? And how does that relate to some of the analyses of rising declaratives you discuss? 



\chapter{Introduction}
\label{chap:introduction}
We use language to perform various kinds of speech acts -- providing information, asking questions, making requests, etc. 

In any language, there are specific signals in the form of a sentence that indicate what speech act it is typically used for. 


In particular, languages tend to have dedicated types of sentences for the same three basic speech acts: declaratives are typically used for assertions (\ref{ex:intro:intro:dec}), interrogatives for questions (\ref{ex:intro:intro:int}), and imperatives for commands (\ref{ex:intro:intro:imp}):

\bex{ex:intro:intro}
\bxl \label{ex:intro:intro:dec}
That's Elmo. \hfill Declarative, Assertion
\ex\label{ex:intro:intro:int} Is that Elmo? \hfill Interrogative, Question
\ex\label{ex:intro:intro:imp} Find Elmo! \hfill Imperative, Request
\exl
\eex


These dedicated types of sentences are commonly referred to as \tit{clause types} (\citealt{sz1985speechact, konig2007, aikhenvald2016, portner2018}, see \cite{konig2020} for a recent review). Different, unrelated languages tend to have the same clause types, but languages differ greatly in how clause types are realized in the surface form. Even if learners have innate knowledge of what the clause type categories \emph{are}, they still need to learn to \emph{associate} the surface expressions of their language with the right clause type. Previous experimental results suggest that by 18 months old, children differentiate these clause types and associate them with their canonical speech act (e.g. \cite{casillas2017turn,marshmallowqueen}). How do children figure it out? %In principle, the association of clause types with particular speech acts should help here. If learners could perceive speech act types directly, this would indeed be a very useful cue to learn how clause types are expressed in their language. But speech acts cannot be perceived directly and, conversely, we often identity a sentence to be performing a particular speech act by the sentence's clause type. There is a potentially vicious circularity here: we typically identify speech acts by using clause type information, but we might need speech act information to learn to recognize clause types. How do learners avoid this circularity?

In the remainder of this chapter, I will clarify how to understand the notions of clause types and speech acts, and discuss some of the specific problems that learners face when acquiring clause types. I will close by outlining the \emph{pragmatic bootstrapping hypothesis} as a potential solution to the puzzle.

%Transition sentence!!!!
In order to get the clearest picture of the puzzle described above, it is useful to start by taking the strongest position about the clause type categories, i.e. that there are three possible values of $C^{0}$. One can then ask how the learner could identify the signals to these categories. Specifically, the learners need to learn how to \tbf{cluster} sentences into these three clause types, and \tbf{label} them with their functions. 


\section{Learning problems related to clause types}


Assuming that the clause type categories are innate, learners only need to identify the signals to the abstract category. Namely, what learners need to figure out is which sentences have the [+int] feature and which ones have [-int]. 
For example, while (\ref{ex:intro:cluster-base}) and (\ref{ex:intro:cluster:dec}) are two different strings, the learner needs to recognize that they share the same feature when it comes to clause types, frequently analyzed as a [+int] feature on $C^{0}$. They also need to recognize that even though (\ref{ex:intro:cluster:dec}) and (\ref{ex:intro:cluster:int}) share the same lexical items, their $C^{0}$ are marked with different features: [+int] for the former, and [$-$int] for the latter.


\bex{ex:intro:cluster-base}
Do you want a cookie?
\eex
\bex{ex:intro:cluster}
\bxl\label{ex:intro:cluster:int}
Is that Elmo?
\ex\label{ex:intro:cluster:dec}
That’s Elmo!
\exl
\eex


Additionally, learners have to learn the language-specific way that [+int] are expressed. A case in point: in English, the subject and verb switch their canonical order in interrogatives (\ref{ex:intro:intro}a-b), but Mandarin employs sentence final particles for interrogatives (\ref{ex:intro:man}a-b). So learners will have to identify which surface morpho-syntactic properties are relevant from the input.  


\bex{ex:intro:man}
\bxl
\gll Zhe shi Elmo.\\
This is Elmo\\
\trans ``This is Elmo." \hfill Declarative
\ex 
\gll Zhe shi Elmo \tbf{ma}?\\
This is Elmo \Sfp\\
\trans ``Is that Elmo?'' \hfill Interrogative
\ex 
\gll Zhizhi Elmo!\\
Point Elmo\\
\trans ``Point at Elmo!'' \hfill Imperative
\exl
\eex


But [\textpm int] and [imp] of matrix clauses are abstract features and cannot be directly read off from the surface string, so the learner is left with the task of figuring out the surface form of each feature. To achieve this, they need to find the right way to cluster the sentences in their language. In other words, when given a new sentence (\ref{ex:intro:cluster-base}), they should be able to put it in the same clause type category as (\ref{ex:intro:cluster:dec}) and not (\ref{ex:intro:cluster:int}). I will refer to this problem as the \tbf{clustering problem}. 

Solving this problem isn't straightforward, as clause type features of $C^{0}$ do not have a one-to-one mapping with morpho-syntactic properties in the surface form of sentences. For example, as discussed above, the [+int] feature of $C^{0}$ in matrix clauses in English is associated with the raising of auxiliaries, and in many cases, resulting in subject-aux inversion. But in embedded clauses, [+int] does not trigger the raising of auxiliaries, and consequently, we won't see subject-auxiliary inversion in these interrogative clauses: %%should mention subject wh

\bex{ex:intro:eng-embed}
Mary wonders \tun{whether Ann can hug Elmo.}
\eex
\bex{ex:intro:eng-can}
\bxl
Ann can hug Elmo.
\ex Can Ann hug Elmo?
\exl
\eex
As shown in (\ref{ex:intro:eng-embed}), the auxiliary \tit{can} and subject \tit{Ann} of the embedded interrogative have the same word order as the matrix declarative sentence (\ref{ex:intro:eng-can}a).


Conversely, some morpho-syntactic properties typically associated with [+int] could also appear in other settings. For example, [+int] in Mandarin can be used to perform a question by having a \twh-phrase in the sentence. But, [$-$int] sentences could also have \twh-phrases, where these phrases are interpreted as indefinites like English \tit{any/a}. As a result, a string like (\ref{ex:intro:m-whamb}) could be either an interrogative (interpretation a) or a declarative (interpretation b). 


\bex{ex:intro:m-whamb}
\gll Xiaoxiao mei 	chi 	\tun{shenme} dongxi\\ 
Xiaoxiao \Neg{} 	eat	what	things\\
a.	``What didn’t Xiaoxiao eat?''	\hfill Interrogative \twh\\
b.	``Xiaoxiao didn’t eat anything.''		\hfill Indefinite \twh
\eex

Therefore, learners need to infer the abstract clause type feature of sentences, but they might not see the surface morpho-syntactic properties associated with clause type feature, or the property that they do see misaligns with the actual clause type feature of the sentence.

Besides the many-to-many mapping problem, learners also have to deal with cases where the relevant morpho-syntactic cues are masked by operations such as ellipsis.  For example, left-edge-ellipsis could obscure morpho-syntactic cues distinguishing declaratives from interrogatives in English (\cite{zwickypullum1983leftedge}):

\bex{ex:intro:lee}
Want to go out?
\eex

The sentence in (\ref{ex:intro:lee}) could be an ellipsis of  the subject pronoun from a declarative like (\ref{ex:intro:lee-unpack}a), or of the subject and the auxiliary from a polar interrogative like (\ref{ex:intro:lee-unpack}b). But the surface form of (\ref{ex:intro:lee}) itself does not have enough information to help us identify its clause type feature. 

\bex{ex:intro:lee-unpack}
\bxl
You want to go out.
\ex Do you want to go out?
\exl
\eex


%

After identifying the clusters, learners need to determine the canonical function of each cluster in the system. That is, after clustering sentences into three categories, children still need to learn which one of these clusters is the interrogatives, which is the declaratives, and which is the imperatives. We will refer to this as the \tbf{labeling problem}. To break into this labeling problem, the learner might need to use the function of the utterances in a cluster -- that is, the speech act information. As many have noticed, clause types are systematically related to the speech act of a sentence (\cite{katzpostal1964, sz1985speechact, portner2018} among many others). If the learner observes that a cluster of sentences is more frequently associated with questions, they might be able to label this cluster of sentences as the one whose canonical function is asking questions (i.e.\ interrogatives), and similarly for declaratives and imperatives. 

But the mapping between speech acts and clause types is not one-to-one either. The function of a sentence can be thought of as the speech act performed by uttering it. For example, Alex's mother uses (\ref{}) to ask Alex's preferences, performing the speech act of questioning.
\bex{}
Do you want to get down? \hfill Alex's mother, Session 01;05;12, Providence Corpus (\cite{ProvidenceCorpus})
\eex

%This paragraph is not good
Speech act theory focuses on what speakers \emph{do} with sentences (\cite{austin1975things, searle1969} a.o.), but this is not the only way to conceptualize speech acts. I will return to the different options for describing speech acts in the next chapter. For now it suffices to observe that, regardless of one's theory of speech act, it is generally agreed upon that the form of a sentence, i.e. its clause type, constrains the types of speech act performed by uttering it. Canonically, declaratives tend to be assertions, interrogatives questions, and imperatives requests. 


But this mapping from clause type to speech act type is not inviolable. In some contexts, it is possible that the conventionalized speech act associated with a clause type is not the actual speech act performed by uttering it. \emph{Indirect speech acts} are these mismatching cases where the primary, ``non-literal'' force of an utterance is different from the conventionalized, ``literal'' force of a sentence associated with its clause type (\citealt{searle1975tax, searle1976class, bachharnish1979, levinson1983, searlevanderveken1985, portner2018, starr2014, murraystarr2020} a.o.). The common example illustrating this phenomenon is \tit{Can you pass the salt?} as a request. As an interrogative sentence, its conventionalized (and ``literal'') force is questioning, but the primary act performed is requesting. As a result, some speech act categories can be expressed by more than one clause types, and vice versa. For example, interrogatives can express assertions, questions, requests/commands (\ref{eng-cl:q-all}); and questions can be expressed by \diis{} (\ref{eng-cl:int-all}).
\bex{eng-cl:int-all}
Speech acts expressed by interrogatives 
\bxl Is it snowing? \hfill Question
\ex Aren't you sweet. \hfill Assertion
\ex Can you pass the salt? \hfill Request
\exl
\eex

\bex{eng-cl:q-all}
Clause types expressing questions
\bxl
Is it snowing? \hfill Interrogative
\ex It's snowing? \hfill Declarative
\ex Tell me if it's snowing! \hfill Imperative
\exl
\eex


\section{Pragmatic bootstrapping hypothesis}
%It is possible that children could use the link between speech act and clause type to make bootstrap into clause type categories

As said in the initial description of the problem, it would be useful for learners to have access to speech act information in order to determine clause type. After separating the problem into the clustering problem and the learning problem, it may even appear \emph{necessary} that they can access speech act information. Surface features alone may allow a learner to cluster sentences into three distinct formal categories, but without being able to determine the conventional function of the sentences in a cluster they will not be able to \emph{label} the clusters correctly. This leads me to postulate the following pragmatic bootstrapping hypothesis:

\begin{quote}
Infants use the speech act information, in addition to observations of morpho-syntactic features in the surface form of sentences, to cluster and label sentences into the three major clause types.
\end{quote}


%previously...here are some of the differences

But, as said, there is a danger of circularity.
Our \hypos{} assumes that children can infer the speech act categories of the sentence at this point. While some evidence suggests that 18-month-olds can infer speech acts, their inferences might not be perfect. Consequently, they might have only limited access to speech act information. Moreover, there's the problem of how children can infer speech act categories in the first place -- and it is undeniable that the clause type information is useful for solving this problem. As adults, the primary way we identify the act performed by a given sentence is through its clause type. But this is precisely the problem that the child is trying to solve (i.e. identifying the clause type). If children need speech act information to identify clause type categories, but they also need clause type information to identify speech act categories, it seems that we have a chicken-and-egg problem. 

We can break out of the circularity by observing that, to get the learning process going, children do not need \emph{perfect} speech act information. I will evaluate the \hypos{} by 
%I address the first problem by
simulating the learning of clause type with various degrees of noise in the speech act information, so that we can see how much pragmatics a learner needs to succeed at the clustering problem. I then tackle the question of where even imperfect speech act information could come from, if no clause type information is accessible,
%address the second problem
by exploring non-clause type cues for speech act information in the input. I find that the prosody of parents' utterances, speech gap between utterances, and direct eye gaze might be useful in learning the distinctions between speech acts.
These are noisy cues for speech act information, but cues nonetheless. As my simulations show that even very noisy speech act information is useful for learning clause types, this suggests a way out of the circularity. Learners can make use of non-clause type cues towards speech act, which may be sufficient to solve the clustering and labelling problems.

\section{Discussion and roadmap}
\label{sec:intro:roadmap} 

This dissertation is organized as follows. Chapter~\ref{chap:background} examines the developmental trajectory of speech acts and clause types, especially questions and interrogatives. As we will see, English-acquiring infants as early as 18 months seem to have already sensitive to the distinctions between different clause types and speech acts, and seem to understand the mapping between questions and interrogatives. The same holds for infants acquiring other languages as well, even though we have less evidence. Our question then is, how do 18-month-olds learn to figure out clause types?

Chapter~\ref{chap:eng-cl} looks at how English-acquiring 18-month-olds could solve the problem. Specifically, is information from syntax enough for children to find the right three clause type categories, or do they need pragmatic information like the speech act of the sentence to find the right clustering? I build two computational models to address this question, a \distlearner{} (\dlearnerabbr{}), and a \praglearner{} (\plearnerabbr{}). These two learners both need to infer the abstract clause type, but \dlearnerabbr{} draws inferences from syntactic information alone while \plearnerabbr{} uses both syntactic and pragmatic information. I use a corpus study to first provide a quantitative description of the type of input that infants get, and use the resulted annotated dataset as input for the computational models. I find that pragmatic information is indeed important for solving the clustering problem: without the speech act information, \dlearnerabbr{} cannot find the right clause types. Additionally, a little pragmatics goes a long way, as  even if 80\% of the pragmatic information is noise, it still improves the learner's performance. 

In Chapter~\ref{chap:man-cl}, I apply the same methodology to another language, Mandarin. Mandarin-acquiring infants figure out the clause types of their language around the same age as English-acquiring infants, but the two languages employ different morpho-syntactic features for clause typing. How do Mandarin-acquiring infants solve the problem? Do they also need pragmatic information? I compare the same two learners, and found that pragmatics information is crucial for identifying Mandarin clause types; without pragmatic information, learners might not be able to identify any of the clause types correctly.

But in both chapters, I operate under the assumption that infants have information about speech act types at their disposal. How do they obtain such information about the speech acts of their parents' utterances? %Of course their knowledge of clause type might help, but are there signals from other sources?
In Chapter~\ref{chap:eng-sp}, I explore potential cues from prosody and parents' behavior that might help infants identify questions. Chapter~\ref{chap:discussion} concludes the dissertation.

