Throughout graduate school, I must have played the scenario of writing this section of the dissertation a thousand times in my head; it always seems so far in the future that it almost seems fictional. I still can’t quite believe that, today is the day that I need to write this part of the dissertation. With sincere apologies to any that I may have forgotten, let me attempt to express my heartfelt thanks to everyone that has taken me to this point.
 
First, my deepest gratitude goes to Valentine Hacquard and Jeff Lidz. Before I even applied to UMD, everyone told me Valentine and Jeff together form a great advising team, and after working with them for five years, I can add my testament to this. Not only did they train me to become a better researcher and a careful thinker, but they were also my cheerleaders. They cheered me on whenever there was any breakthrough, no matter how small it was. I cannot count how many times I was about to give up on an idea, and only pushed through because Valentine and Jeff said “what are you talking about, this is great!” I feel so fortunate to have them as my mentors.
 
Thanks to my committee members, Naomi Feldman, Daniel Goodhue, Alexander Williams, and Yi Ting Huang. I was so nervous when I reached out to Naomi to ask for her help with building a computational model, because I barely understood anything from her class on cognitive modeling. But it turned out that I had nothing to worry about. Naomi was so kind and patient, even when she had to explain to me the extremely basic math for the nth time. Without her, I wouldn’t even dare to ask the ``how” question of learning, and half this dissertation wouldn’t exist. She also connected me with people at the CLIP lab, which saved me so much time. Dan has always been a great mentor, and I had the fortune to both be a mentee, and a co-mentor. I learned so much about mentorship when we worked with a group of undergraduate students together. He always gave the best advice: just challenging enough to make the work better, but not so challenging that you feel defeated. He also always made sure to add some encouragement in his comments, and I’m very grateful for that. Many thanks to Alexander for demonstrating to me how to be a careful thinker. He’d offer the most thoughtful and eloquent comments, often end with a joke that I fail to get (the Chomsky-Wittgenstein-Slab joke still puzzles me). I always wished I had recorded his comments at meetings, so I can write them down verbatim. Being his TA for his Philosophy of Language class was one of the most enlightening experiences for me as a semanticist.
 
I feel so lucky to have spent five years in such an amazing department. This community makes linguistics so much fun, and you know you are in the right place when you look forward to going to work every day. Thank you to Howard Lasnik, for reminding me how breathtakingly beautiful syntax can be, and for giving me the only A+ I’ve ever gotten; to Omer Preminger and Norbert Hornstein, for showing me how to think outside the box, and for reminding me to never lose sight of the big picture; to Colin Philips, for teaching me not to do experiments when you don’t have to, and do it carefully when you do have to; to Ellen Lau and Philip Resnik, for demonstrating to me how to make other cognitive scientists care about linguistics, and how to make linguists care about other fields of cognitive science; to Peggy Antonisse and Tonia Bleam, for teaching me how to teach in America; to Shevaun Lewis, for giving me so much good advice over the years and for puzzling over the PuffinMan data with me; and of course, to the incredible Kim Kwok, for always being so patient with me when I made mistakes filling out forms and for teasing me for making them. I really enjoyed our daily chats at 9 in the morning that one year; too bad it stopped because of COVID.
 
Great teachers and mentors are hard to find, but I somehow found many over the years. I feel so unbelievably fortunate that I met Yang Xiaolu, Yang laoshi, and Roger Oleson in Tsinghua. I took my first acquisition class with Yang Xiaolu, \tit{Yang laoshi}, and got to follow her around at conferences and lab meetings during my senior year in college. She always encouraged me to ask questions and present my ideas. Looking back, some of my ideas were laughable. But Yang laoshi never laughed at them, she just helped me make them better. Even with all her guidance, she still made me feel that I had ownership over anything I was working on. She also cared about her students. When I was having a hard time after graduating college, she was there to offer comfort and advice. Thank you, Yang laoshi! 

I couldn’t remember which class I took with Roger, but it’s not important. In fact, I never thought of him as a teacher (because teachers are on a higher social hierarchy in my mind) but as a friend. I want to thank him for always reminding me that the important things in life are not your job, but your passion, your friends and family, and maybe Sichuan food.
 
Before UMD, I had the opportunity to spend a semester at UMass, and Angelika Krazter and Tom Roeper made this semester so special. Angelika opened the door of semantics for me. Still today when I run into a semantics problem, I always go back to her notes, and try to think like Angelika. 

I didn’t expect it, but Tom’s class, and our regular weekly chat quickly became my favorite time of the week at UMass. Brainstorming with Tom was the best. Before I came to UMass, I had my creativeness and passion beaten out of me, and was about to give up linguistics. Tom helped me find my passion and imagination back. Nothing was too crazy for Tom, and he never looked down on me when I presented him with a crazy idea. I can’t thank him enough for making linguistics enjoyable for me again. He was also extremely kind, and it has always been so great to run into him at acquisition conferences.
 
I only took one syntax class with Li Yafei, and one with James C-T Huang, but these two classes were so memorable and inspiring that I still think about them today. Before their classes, I didn’t really enjoy syntax, because I thought it was just about memorizing the right generalizations (yeah, horrifying, right?). They completely threw that idea out of the window. For the first time, I learned about \tit{how} to do syntax, and it has NOTHING to do with memorization. Thank you, Li laoshi, Huang laoshi!%Even though this dissertation is not a syntax dissertation, but without their \tit{dianhua} (I’m really not sure which English word would be the best equivalent, perhaps ``enlightenment”), I wouldn’t even be able to ask the questions I asked.

Thank you to Daniel Altshuler. I didn’t know it at the time I met Daniel, but he was about to change my life. I still don’t understand why he decided to help me, but it might be just because Daniel is such a wonderful person. Thank you for looking out for me and Julian over the years, and can’t wait to see you again at conferences/summer schools!
 
I once heard that “Let’s work on this together!” is the academics’ way of saying “let’s start a band!” I’m so glad to have started bands with so many people over the years. Thank you to Mingming Liu, for teaching me about alternatives and exhausitivity so many times, and for tolerating my crazy questions. Working with you was very rewarding.  Thomas Schatz makes ASR less intimidating, and no matter what problem you want to solve, Thomas had written a script for it. Thank you for making it possible for me to have something related to prosody in this dissertation, I got so tired of people asking me about it and it felt really good that I can point them to this chapter. I’m really looking forward to continuing working on it to make it better. Nick Huang, you have always been my role model. Your work ethic is unparalleled, and you were always so thoughtful. It was so wonderful when you were still in the department to be able to just pop in your office and talk about linguistics, and I’m so glad we decided to officially work together last year. I hope we will continue our collaboration! And on that note, thank for introducing me to Aaron White and Zhendong Liu, and hope our project together will turn into a paper soon. Thank you to Tyler Knowlton. We tried to work together so many times but our experiments just keep getting mediocre results. I really hope that one day one of our ideas will work! Thank you to Liao ChiaHsuan, hope we’ll pick up the \twh-indefinite project again someday! Liu Ying, you were my first non-supervisor collaborator, and we had to teach ourselves everything from lambda calculus to focus semantics, and it was the most rewarding experience. Gaor, I hope we get to work together again!
 
The acquisition lab at UMD is a special place. Jeff always said it should be as professional as a dentist's office, but also as fun as a kindergarten (or preschool?). Indeed, I learned how to be a professional, and I had a lot of fun. The lab meeting was my favorite place to present my work, and I always walked out of 1108B with a better project. Thank you to my lab mates over the years: Laurel Perkins, Nick Huang, Mina Hirzel, Tyler Knowlton, Hisao Kurokami, Adam Liter, Iman Bou-Saboun, Jack Ying, and Katherine Howitt, for all the stimulating questions and great suggestions; and of course to Tara Mease, for teaching me how to be a good project manager. I also want to thank all of my RAs over the years: James Burns, Xiaoyu Yang, Ziqing Ji, Rin Gourianova, Luke Burger, Avni Gulrajani, Jae Yu, Eli Herbst, and Zhang Jiaobao, for putting up with tedious annotation work. I also got to work with James Burns, Xiaoyu Yang, Ziqing Ji, and Avni Gulrajani more closely in the past year, and it was such a wonderful experience. Thank you!
 
My sincere thanks to everyone at Haidian Xibeiwang Xuequ for helping me run my studies in Beijing from 2017-2019.  I want to especially thank Chen Shulan, Miao Zhibin, Zhang Xuening, and Zhao Xiaojing for their kind accommodation.
 
I’ve said this to so many people, but office 3416F is magical. Over the years, I’ve had the best officemates: Anouk Dieuleveut, Aaron Doliana, J\'essica Mendes, and Mina Hirzel. I don’t have any siblings, but you guys are like my sisters and brother.
 
Anouk, what can I say. You are like a sister to me. Together, we laughed, cried, and shared so many hugs and so much bread. It was so hard for me to say goodbye to you last year; just writing this filled my eyes with tears. I’m so glad we decided to keep working together via zoom, and it always makes me so happy to see your face. Thank you for everything!
 
Aaron, thank you for being my friend. You were the first person I met at UMD, and I’m so happy we got to share an office for three years (well, one and a half really because of COVID). Our “reading and Chinese takeout group of two” was indeed a good time. Thank you for always reminding me that you don’t have to stress yourself out as a graduate student. Thank you also for recruiting Divya and Myra for our hot pot party after Tyler and Zoe dropped out.

 
J\'essica, when I heard that you decided to come to Maryland, I was so happy. I almost immediately requested to have you as an officemate, but unfortunately, COVID delayed that for almost a year and a half. You made me laugh so hard. Thank you for being such a great friend, and sorry I wasn’t so much fun when I started writing this dissertation. I hope we can keep in touch!
 
Mina, you are simply amazing. I admire you as a researcher and love you as a friend. Whenever I need to present something, I always ask myself, how would Mina do it? Your presentations and posters were always so clear, so beautiful, and so fun. But you are also so generous with your time and gave me so much great advice over the years. Hanging out with you (and occasionally Jad) every Sunday in your beautiful garden during COVID was the highlight of my week that year. Thank you for everything, and I can’t wait to go to all the good restaurants with you in Boston!
 
Many thanks to my cohort, Adam Liter, Hisao Kurokami, Jackie Nelligan, and Hanna Muller, and to Laura, Nima, Jon, and Aura. We had many good memories sharing an office in our first year, and later hanging out in Adam’s backyard. Laura, I’m so looking forward to going to karaoke with you and Katherine again!
 
I also want to thank my friends at UMD over the years whom I hadn’t shared an office with (but I wish I had): Paulina Lyskawa, Gesoel Mendes, Nick Huang, Ted Levin, Phoebe Gaston, Kasia Hitczenko, Laurel Perkins, Annemarie von Dooren, Tyler Knowlton, Laurel Whitfield, Jad Wehbe, Clara Cuonzoe, Luisa Seigun, Craig Thorburn, Allyson Ettinger, Anton Malko, Sigwan Thivierge, Masato Nakamura, Alex Krauska, Polina Pleshak, Joselyn Rodriguez, Nika Jurov, Lesli Li, Rosa Lee, Jack Ying, Yichi Xu, Chen Jingyi, Katherine Howitt, Zulfiyya Aghakishiyeva, London Dixon.
 
Thank you, Tyler and Zoe, for hosting so many hot pot parties. They were some of the best memories I had in Maryland. Remember that one time we had to wait until 11pm for the delivery of Asian groceries? It was definitely worth it. Also, thank you both for always answering my questions about cognition, and Tyler, hopefully one day one of our many efforts to work together will work out! Paulina, thank you for always checking in on me, it really made a difference. Annemarie, you have the best style, and the kindest smile, and hope we’ll get to see each other again sometime! Craig, thank you for helping me navigate the server, and for forgiving me when I accidentally flooded the \textsf{realspeech} directory (sorry again everyone). Anton, thanks for making R a little bit more understandable. Alex, I can’t believe we only found out about our shared love of the violin three months before I defend, and only got to play together once after my defense! But it was so much fun, and hope we will get to play together again sometime. Sig, you pulled the best prank, kudos! It still makes me laugh. Clara, thank you for always stopping by my office and giving me a hug! Allyson, sorry about bugging you about \tit{ba}, I didn’t understand how stressful you must have been; thank you for being so patient. Laurel, your work never fails to inspire me, and thank you for answering my many questions about them!
 
Thank you to my roommates in Maryland: Hallie Oines, Rana Karimpour, Beril Yalcinyaka, Julia Brown, and Erika Dömötör. I’ve been so lucky to have found the best roommates. Thank you for being so kind and understanding. Thanks to Zoe Schluter for introducing me to the house. Thanks to Patrick and Koki Smith for letting me live in their house for five years.
 
Thank you to Niki, Kelley, and Kathy at A Cat’s Life Rescue. Niki, I can’t imagine doing what you do. You are always so patient with my questions (most of the time the answer was “that’s normal”), always have the next kitty for me to foster, and never said no when I asked you to help a kitty. Thank you so much for coming to my house at 2am when Ally passed away to give me a hug. Seeing you and Phoenix always makes me happy. Kelley and Kathy, thank you for giving me some of the best foster babies (I miss Boss and Milo so much), and for always being there with medication and supplies when I need them. I couldn’t stop laughing because half of them have the word ``poop”, and the other half are adorable pictures of said poop (for those of you who are lucky enough to have never worried about cat poop: soft poop might be a sign for many health issues for kitties). I’m so grateful that all of you at ACLR gave kitties with all kinds of pooping- and non-pooping-related issues a chance.
 
I feel so lucky that I got into linguistics, because it’s such a great community. I made so many friends in this community, which makes going to conferences so much fun. Thanks to Rob Pasternak, Leah Chapman, Kimberly Johnson, Andrew Lamont, Zahra Mirrazi, Alex Göbel and Emma Nguyen-Göbel, Nathan Huang, Tom Maxfield, John Ander Mendia, Deniz Özyildiz, Mike Clauss, for making my semester at UMass feel so welcome; thanks to Jess Law, Haoze Li, Yi-Hsun Chen, Mingming Liu, Beibei Xu, Hazel Mitchley, Shuhao Shih for a great time at Rutgers, and for giving my first conference presentation after leaving Hong Kong less nerve-wracking; Thanks to Maxime Tulling, Daniel Altshuler, Yimei Xiang, Lucas Champollion, Ailis Cournane, Vera Zu, Wu Zhuang, Aaron White, Rachel Dudley, He Yuyin, Sun Yenan, Jackie Lai, Sherry Chen Yong, Xu Ting, Ji Yue, for all the great conversations and advice at conferences, and sometimes for hosting me when I couldn’t afford hotels. The Barcelona ESSLLI connected me with so many wonderful people (including my future spouse), but a special thank you to Jenny Tan, Matt DeVilbiss, Markus Brenner, Anya Zaretskaya, Adam Kupś, Kuba Kozakoszczak, Lina Brixey. I will never forget all the delicious tapas, the beautiful breeze, and all the walks around Barcelona. Thank you my friends for reminding me how wonderful life could be.

 
I met some of my best friends in Hong Kong. Liu Ying, Jia He, Jess Law, Haoze Li, Ai Shu, Jia Li, Qun Li, Shi Xinyue, Guo Li, Yvonne Lee, Eunice Yip, and Xiaochun Hong, I cannot say thank you enough. I wouldn’t be able to survive those years without you guys.
 
Liu Ying, aka Gaor, thank you for everything. We were both in a tough situation, but you still decided to help me. You gave me hope and courage when I needed them the most. I hope you know how much your friendship means to me, and I don’t know how I can ever repay you. I miss you deeply, and hope I can see you and your baby soon!
 
Haoze and Jess, I don’t even know where to begin. When I felt like all the roads were blocked, you guys helped me find a way. I admire you both, and am so proud to call you my friends. I hope one day we get to have dinner again, and maybe Haoze’s cooking will be even better thne!
 
He jia jie, you were the first one in Hong Kong who asked me how I was doing. For that and for so many other things, I am forever grateful. You have always been so fun to hang out with, and I’m so sorry we fell out of touch over the years. I really hope I get to see you again soon!
 
Ai Shu, Qunzi, and 2 Jia, thank you for always lending me sympathetic ears, and for tolerating me when I was being a selfish brat. I miss you guys, and hope we’ll get to \tit{dabianlu} again one day.

Li Guo, our bi-weekly syntax/prosody reading group was a lot of fun! Your knowledge of Classical Chinese is impressive. Thank you for always answering my questions about the grammar of Classical Chinese.
 
Yvonne and Eunice, thank you for noticing that something was wrong. Special thank you to Eunice, for telling me leaving doesn’t mean anything, but just that this place isn’t right for me. If anyone is reading this dissertation, I want to pass that advice to you. Sometimes it might feel like there is no way out, but it's almost never true. 
 
Thank you to my friends at Tsinghua: Shi Peipei, Qian Chen, Guo Jiabao (Carbo Kuo), Zhou Dong, Cao Tianchen, Zhu Qianyun, Li Meirong, Qian Yuzhu, Huang Binhuan, Lei Zhongxing, Zhang Wenxiu, Roger Olesen, Gwyneth Ho. Gwyneth, I’ve always admired your courage and determination. I hope one day I can hear your stories directly from you again, and not from the news. Thank you to Yang Nan, Shi Hanning, Yu Ning, for being such great friends.

Thank you to my extended family in Chengde. Yeye, Laolao, Laoye, I miss you all. It still hurts that Nainai passed away while I was in America. I know she loved me, and would be proud of me. Thank you to Laoyi and Laoyifu, Dagu, Ergu and Ergufu, Laogu, Jiujiu and Jiuma for your support; thank you to my cousins Zhutianjie, Jin'aoge, Xinhe, Yang Rousi for your friendship (with lots of teasing of course). Because of graduate school, I missed so many important family moments, and I can only hope I can see you all soon. 

Thank you to my in-laws Anna and Bernd Schl\"oder, and my sister- and brother-in law Nikki, Lotte and Domi. I think about how lucky I am to have you guys all the time. Thank you for letting me play your 250-year-old violin! It's the best Christmas-birthday-graduation gift ever. Hope to spend Christmas with you soon!
 
Everyone says this, but I really do have the best parents. Thank you for everything, Chang Yang and Dr.~Jun Liu! Ma, you are such an inspiration to me. You never back down from a difficult situation, and I hope you know how proud I am being your daughter. I want to tell everyone: “you see that fiercely brilliant woman, that’s my mom!” When my friends in the US told me the word “doctor” usually makes people think ``male,'' I laughed so hard because the best doctor I know is you. I learned how to be independent, thoughtful, and kind from you, and I hope you are ok with having another doctor in the family! Ba, sorry for always running away when you want to teach me something, but I hope you know how much I admire and love you. I didn’t know how valuable it is to have such loving and understanding parents, until I saw how much my friends suffer because of their parents. I miss you both so much, and I hope I get to say \tit{wo ai nimen} in person soon.
 
Julian, thank you for sticking with me through graduate school. There have been so many ups and downs (writing being a huge down), but you are always there for me, sometimes with a funny story to make me laugh. I really couldn't have done it without you. Thank you for being such a great partner. The future may seem terrifying at times, but as our officiant said, we’ll “let love lead the way.” I love you and I like you.
 
 

